### 2022-03-24 abstract.tex
review of ch 1

This dissertation focuses on how people search for information and how people
rely on this information to inform their health behaviors and develop social
norms. Information can be shared and promoted by others
through information campaigns and consumed by an individual, called information
push, or information can be intentionally sought out by individuals, 
called information pull, and how the information then relates to the
creation of social norms. In academia and policy, the focus in research on 
information has most often focused on the information that is sent to 
consumers, seeing people as passive receivers of information.
My dissertation instead focuses on individuals as active agents 
in their search for information.

The first article tests the construct validity of the Google Search Trends source,
an online indicator of computer-mediated information search. I attempt to
validate Google Search Trends for use as indicators of attitudes, disease prevalence
and political preferences using five different validated data sources. 
I fail to find correlation among any of the Google Trends tested and their
validated indicators. I demonstrate that there is no criterion validity of
Google Trends for the selected cases and social scientists will find no replacement for
high quality survey data with Google Trends. Instead, we must only use 
Google Trends to demonstrate interest or attention.

Knowing that Google Trends data only encompasses a small portion of the
information-seeking done by modern humans, my second article 
is motivated by the research question: How do computer-mediated or interpersonal 
information-seeking strategies vary across populations? Using original survey data
of 948 Americans, I investigate their experiences seeking out information about 
the COVID-19 vaccines among the following information seeking modalities: 
personal connection, doctor, social networking site, online forum, and online search engine. 
I find little evidence that online search is more utilized than
seeking social support from personal network connections or health professionals
as I had hypothesized based on uses and gratifications theory, 
that media users are not passive consumers but instead actively chose
the media sources based on the satisfaction of their individual needs.
I find that different exposure points, the ways people first are exposed to 
information without searching for it, and information search modalities, the 
ways people intentionally search for information, hold real world consequences
through their associations with COVID-19 vaccination rates and intentions.
For example, receiving or seeking out information from a doctor increases 
 COVID-19 vaccination uptake while receiving information from a social 
 networking site like Facebook or Twitter was associated with lower odds of vaccination.

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty, specifically using the cases of
 stay-at-home rates and vaccination rates as responses to public health
recommendations to mitigate the COVID-19 pandemic. Using theories like associative diffusion
and the integrated theoretical framework of norms, I test models of behavioral adaption
to public health recommendations and patterns of complex contagion (the need for
repeated exposures to a something novel for it to diffuse) using linear
mixed effects models. My results show that complex contagion is a valid 
framework for the social contagion of new norms during COVID-19. 
However, I find an important novel moderating effect of signal discordance; 
if there is diversity in the information received by an ego, 
contagion is less likely to occur. This paper shows that
the contagion process is not fully understood without looking at the context
of each exposure to a contagion within the range of contagions one experiences.

The introductory chapter provides a summary of the research and an explanation of how this
research contributes to the sociology, social science, and society. Namely, 
this paper provides important perspectives on search as an agentic process
and how the micro-level information seeking process of an individual can
lead to macro-level social norms. I show that information diffusion
is disrupted when conflicting information and behaviors are simultaneously
diffusing and therein contribute to research on diffusion, social networks,
and social norms.

>>>>

This dissertation focuses on how people search for information and how people
rely on this information to inform their health behaviors and develop social
norms. In academia and policy, the focus in research on 
information has most often focused on the information that is sent to 
consumers, seeing people as passive receivers of information. This
is otherwise known as a focus on the information 'push'. But, 
information is also intentionally sought out by individuals;
called information pull. My dissertation focuses on individuals as active agents 
in their search for information and how the information discovered through
these push and pull processes relates to the creation of social norms.

My first article tests the criterion validity of Google Search Trends as
an indicator of computer-mediated information search. I attempt to
validate Google Search Trends for use as indicators of attitudes, disease prevalence
and political preferences using five different validated data sources. 
My analysis revealed no correlation among any of the Google Trends tested and their
validated indicators. I demonstrate that there is no criterion validity of
Google Trends for the selected cases and social scientists will find no replacement for
high quality survey data with Google Trends. Instead, we must only use 
Google Trends to demonstrate interest or attention.

Knowing that Google Trends data only encompasses a small portion of the
information-seeking done by modern humans, my second article 
is motivated by the research question: How do computer-mediated or interpersonal 
information-seeking strategies vary across populations? Using original survey data
of 948 Americans, I investigate their experiences seeking out information about 
the COVID-19 vaccines. I investigate five distinct information seeking modalities, or methods of searching for information: 
personal connection, doctor, social networking site, online forum, and online search engine. 
% I find little evidence that online search is more utilized than
% seeking social support from personal network connections or health professionals
% as I had hypothesized based on uses and gratifications theory, 
% that media users are not passive consumers but instead actively choose
% media sources based on the satisfaction of their individual needs.
I find that different exposure points, the ways people first are exposed to 
information without searching for it, and information search modalities, the 
ways people intentionally search for information, hold real world consequences
through their associations with COVID-19 vaccination intentions and rates.
For example, receiving or seeking out information from a doctor increases 
 COVID-19 vaccination uptake while receiving information from a social 
 networking site was associated with lower odds of vaccination.

My final article takes a deeper dive into the formation of social norms
governing health behaviors in cases of extreme uncertainty, specifically using the cases of
 stay-at-home rates and vaccination rates as responses to public health
recommendations to mitigate the COVID-19 pandemic. Using the theories of associative diffusion
and the integrated theoretical framework of norms, I test models of behavioral adaption
to public health recommendations and patterns of complex contagion (the need for
repeated exposures to a something novel for it to diffuse) using linear
mixed effects models. My results show that complex contagion is a valid 
framework for the social contagion of new norms during COVID-19. 
However, I find an important novel moderating effect of signal discordance; 
if there is diversity in the information received by an ego, 
contagion is less likely to occur. This paper shows that
the contagion process is not fully understood without looking at the context
of each exposure to a contagion within the range of contagions one experiences.

The introductory chapter provides a summary of the research and an explanation of how this
research contributes to sociology, social science, and society. Namely, 
this paper provides important perspectives on search as an agentic process
and how the micro-level information seeking process of an individual can
lead to macro-level social norms. I show that information diffusion
is disrupted when conflicting information and behaviors are simultaneously
diffusing and therein contribute to research on diffusion, social networks,
and social norms.

### 2022-03-24 ch1.tex
review of ch 1

The information that spread to communities in Arizona contributed to the health
behaviors exhibited today by Arizonans. While the information spread in various
ways, including by means of diffusion and polarization, it contributed to our newly
established norms in the Spring of 2020 and since. By uncovering how norms were
created during this time of norm uncertainty, my research will contribute to the
recommended interventions to promote effective health-related behavioral norms
before politicization and misinformation occurs. This will create a healthier,
happier Arizona.

@@

Some communication theorists ask why an answer to a question is sought in the
first place. The Theory of Uncertainty Management
\citep{brashersCommunicationUncertaintyManagement2001} professes that people
search for information when their uncertainty around the subject leads to
anxiety or other cognitive harms. The Theory of Motivated Information Management
\citep{afifiSeekingInformationSexual2006, afifiTheoryMotivatedInformation2004}
extends the prior by adding that uncertainty itself is not the catalyst for
information-search; rather, it is driven by a discrepancy between the current
level of uncertainty on a subject and desired level of uncertainty.

@@

Adapting UGT to my own purposes, I theorize that
informational support will be activated from network ties when the informational
need is accompanied by or related to a need to form or maintain social ties
\citep{baumeister,grieve2013face}. For instance, Dani, a young college student, 
could search on Google to find how to turn iron their dress shirt properly, 
but instead Dani choses to call their parent(s) and ask the best strategy so their parent(s)
feels respected and valued and their relationship is strengthened. 
Conversely, I theorize that search will be conducted outside of network
activation when information is needed but an individual does not have 
an additional cognitive need to fill through social interaction. 
For example, Taylor could ask their neighbor who is a medical professional about 
the newfound headaches they've been experiencing, but every time they talk to their
neighbor they ask for help cleaning their gutters, which is too much work; Taylor
instead may choose to ask the Facebook group they're a part of to gain information.
Alternative goals such as identity management or relational maintenance
\citep{brashersInformationSeekingAvoiding2002} need to be balanced and may also
influence where search is conducted.

@@

Social norms form the building blocks of social organization and have been a
focus of sociology since the beginning. In a recent Annual Review of Sociology
piece focused on social norms, Horne and Mollborn define norms as "group-level
evaluations of behavior ... [or] when people have expectations about how others
evaluate behaviors" \citeyearpar[p. 468-69]{horneNormsIntegratedFramework2020}.
One of the principal architects of sociology, \'{E}mile Durkheim, saw that
society exerted powerful forces on individuals and classified people's norms,
beliefs, and values as parts of collective consciousness that provide for
societal integration \citeyearpar{durkheimSuicide1897, durkheimDivisionLaborSociety1933}.
For Durkheim, norms were therefore the glue that held society together in cohesion.

@@

To investigate information-seeking behaviors, online search tools, and how
new norms in health behaviors form, I first provide an outline of the 
information push and pull process in Figure \ref{fig:concept-map}. After 
being exposed to new information, an actor can choose to further investigate
the information online or offline and, in turn, further share the 
information they were exposed to.

@@

I fail to find correlation among any of the Google Trends tested and their
validated indicators. While some Google Trends tested were significantly
associated with the outcome in the regression models, effects tended to be small
and the total model interpretability remained low, even when controlling for
demographic variables. This article shows that there is no external or 
criterion validity of Google Trends for these uses and social scientists will find no replacement for
high quality survey data with Google Trends. Instead, we must only use 
Google Trends to demonstrate interest or attention.

@@

These models show that complex contagion is a valid 
framework for the social contagion of new norms during COVID-19. Importantly, I find
a novel moderating effect of signal discordance, the contextual diversity of
signals received by an ego. If there is diversity in the information received
by an ego, contagion is less likely to occur. This paper shows that the contagion 
process cannot be not fully understood without looking at the context of each
exposure to a contagion within  within the range of contagions one experiences.

@@

% search as an agentic process
This research makes contributions to multiple subfields and literatures. 
First and foremost, these articles contribute to the sociological literature.
The spread of information has long been investigated in the discipline (diffusion, mass
communication, social influence). However, it is almost always researched from a structural
perspective which disregards individual agency and choice in the
acquisition of new information. This dissertation adopts the second perspective,
centering people as active agents in their search for information aligned
with their own preferences and gratifications instead of just
passive receivers of information signals.
Taking this agentic perspective is critical in the study of information
diffusion when combined with cognitive perspectives like that of
\citet{goldbergSocialContagionAssociative2018}. How people
search for information is a strong determinant of 
the information that they find. The information they uncover is filtered
through cognitive biases and predispositions to how they interpret and
act upon the newfound information. Searching for informational support
through social network ties theoretically will uncover potentially
different information than would be uncovered through online web
search and will change how people are organized and how people behave.
\hyperlink{paper-2}{Article 2} Article 2, for instance, demonstrates that 
the choice to receive a COVID-19 vaccine is impacted by how
individuals search for information on the vaccine. 
This is a real-world consequence to the variation in the individual information
search process.

% CSS & big data
This dissertation also contributes to computational social science \&
critical big data studies. I base much of my dissertation around Google
Search Trends, which are relatively underutilized in the social sciences
compared to health sciences and business. Because \hyperlink{paper-1}{Article 1} 
takes a critical big data studies approach, I consider the potential
pitfalls that any source of big data face \citep{mcfarlandBigDataDanger2015} and explicitly
test just how well Google Trends can be used as an indicator for different concepts. 
Importantly, I find that there is no replacement for high quality survey data with Google Trends. 
This is a large contribution to the field of critical big data studies but also provides
the groundwork for future tests of external validity for big data for future computational
social scientists.

@@

% social norms
Additionally, this dissertation investigates how social norms are
formed in situations of uncertainty, an area that is far less developed in
sociology. While sociologists have studied social norms since the
formation of the discipline, norms are most often studied under the lens
of the emergence of norms and transmission of norms between groups, 
deviance and social control, the interaction of norms as "mental 
representations of appropriate behavior" \citep{aarts2003silence}, 
and research categorizing different types of norms (e.g., descriptive 
versus injunctive, prescriptive versus proscriptive).  
The focus on the formation of new social norms during times of upheaval 
has implications for interventions in social behavioral outcomes, 
research on cultural change, and the spread of cognitive associations. 
Moreover, \hyperlink{paper-3}{Article 3} focuses on how conflicting
behaviors become settled into cognitive associations, providing
a relevant framework for research on unrest and disasters.

@@

% social epidemiology
I also make important contributions to social epidemiology through this
research. Each of the three papers in this dissertation use health case
studies for questions of sociological interest. Because of this, I 
contribute to the social epidemiological conversation on vaccine hesitancy,
health communication, and diffusion of high-risk health behaviors. While
all these areas are important to social epidemiology, there hold
potential important health implications for these contributions on
outside of the academy, covered in the first paragraphs of this chapter.

>>>>

The information that spread to communities in Arizona contributed to the health
behaviors exhibited today by Arizonans. While the information spread in various
ways, including by means of diffusion and polarization, it contributed to our newly
established norms since the Spring of 2020. By uncovering how norms were
created during this time of norm uncertainty, my research will contribute to the
recommended interventions to promote effective health-related behavioral norms
before politicization and misinformation occurs. This will create a healthier,
happier Arizona.

@@

Some communication theorists ask why an answer to a question is sought in the
first place. The Theory of Uncertainty Management
\citep{brashersCommunicationUncertaintyManagement2001} professes that people
search for information when their uncertainty around the subject leads to
anxiety or other cognitive harms. The Theory of Motivated Information Management
\citep{afifiSeekingInformationSexual2006, afifiTheoryMotivatedInformation2004}
extends the prior by adding that uncertainty itself is not the catalyst for
information-search; rather, it is driven by a discrepancy between the current
level of uncertainty on a subject and desired level of uncertainty.
% connect better to social support if possible

@@

Adapting UGT to my own purposes, I theorize that
informational support will be activated from network ties when the informational
need is accompanied by or related to a need to form or maintain social ties
\citep{baumeister,grieve2013face}. For instance, Dani, a young college student, 
could search on Google to find how to iron their dress shirt properly, 
but instead Dani choses to call their parent(s) and ask the best strategy so their parent(s)
feels respected and valued and their relationship is strengthened. 
Conversely, I theorize that search will be conducted outside of network
activation when information is needed but an individual does not have 
an additional cognitive need to fill through social interaction. 
For example, Taylor could ask their neighbor, who is a medical professional about 
the newfound headaches they've been experiencing, but every time they talk to their
neighbor they ask for help cleaning their gutters, which is too much work; Taylor
instead may choose to ask the Facebook group they're a part of to gain information.
Alternative goals such as identity management or relational maintenance
\citep{brashersInformationSeekingAvoiding2002} need to be balanced and may also
influence where search is conducted.

@@

Social norms form the building blocks of social organization and have been a
focus of sociology since the beginning. In a 2020 Annual Review of Sociology
piece focused on social norms, Horne and Mollborn define norms as "group-level
evaluations of behavior ... [or] when people have expectations about how others
evaluate behaviors" \citeyearpar[p. 468-69]{horneNormsIntegratedFramework2020}.
One of the principal architects of sociology, \'{E}mile Durkheim, saw that
society exerted powerful forces on individuals and classified people's norms,
beliefs, and values as parts of collective consciousness that provide for
societal integration \citeyearpar{durkheimSuicide1897, durkheimDivisionLaborSociety1933}.
For Durkheim, norms were therefore the glue that held society together in cohesion.

@@

To investigate information-seeking behaviors, online search tools, and how
new norms in health behaviors form, I first provide an outline of the 
information push and pull process in Figure \ref{fig:concept-map}. After 
being exposed to new information, an actor can choose to further investigate
the information if their level of uncertainty on the subject surpasses
the desired level of uncertainty \citep{brashersCommunicationUncertaintyManagement2001}. 
Search can then be performed online or offline and be pushed further to the actor's 
networks.

@@

I find no correlation among any of the Google Trends tested and their
validated indicators. While some Google Trends tested were significantly
associated with the outcome in the regression models, effects tended to be small
and the total model interpretability remained low, even when controlling for
demographic variables. This article shows that there is no external or 
criterion validity of Google Trends for these uses and social scientists will find no replacement for
high quality survey data with Google Trends. Instead, we must only use 
Google Trends to demonstrate interest or attention.

@@

These models show that complex contagion is a valid 
framework for the social contagion of new norms during COVID-19. Importantly, I find
a novel moderating effect of signal discordance, the contextual diversity of
signals received by an ego. If there is diversity in the information received
by an ego, contagion is less likely to occur. This paper shows that the contagion 
process cannot be not fully understood without looking at the context of each
exposure to a contagion within the range of contagions one experiences.

@@

% search as an agentic process
This research makes contributions to multiple subfields and literatures. 
First and foremost, these articles contribute to the sociological literature.
The spread of information has long been investigated in the discipline (diffusion, mass
communication, social influence). However, it is almost always researched from a structural
perspective which disregards individual agency and choice in the
acquisition of new information. This dissertation adopts the second perspective,
centering people as active agents in their search for information aligned
with their own preferences and gratifications instead of just
passive receivers of information signals.
Taking this agentic perspective is critical in the study of information
diffusion when combined with cognitive perspectives like that of
\citet{goldbergSocialContagionAssociative2018}. How people
search for information is a strong determinant of 
the information that they find. The information they uncover is filtered
through cognitive biases and predispositions to how they interpret and
act upon the newfound information. Searching for informational support
through social network ties theoretically will uncover potentially
different information than would be found through online web
search and will change how people are organized and how people behave.
\hyperlink{paper-2}{Article 2}, for instance, demonstrates that 
the choice to receive a COVID-19 vaccine is impacted by how
individuals search for information on the vaccine. 
This is a real-world consequence to the variation in the individual information
search process.

% CSS & big data
This dissertation also contributes to computational social science \&
critical big data studies. In two articles, I employ Google Search Trends 
which are relatively underutilized in the social sciences
compared to health sciences and business. Because \hyperlink{paper-1}{Article 1} 
takes a critical big data studies approach, I consider the potential
pitfalls that any source of big data face \citep{mcfarlandBigDataDanger2015} and explicitly
test just how well Google Trends can be used as an indicator for different concepts through
criterion validity. Importantly, I find that Google Trends cannot replace
high quality survey data. 
This is a large contribution to the field of critical big data studies but also provides
the groundwork for future tests of external validity for big data for future computational
social scientists.

@@

% social norms
Additionally, this dissertation investigates how social norms are
formed in situations of uncertainty, an area that is far less developed in
sociology. While sociologists have studied social norms since the
formation of the discipline, norms are most often studied under the lens
of the emergence of norms and transmission of norms between groups, 
deviance and social control, the interaction of norms as "mental 
representations of appropriate behavior" \citep{aarts2003silence}, 
and research categorizing different types of norms (e.g., descriptive 
versus injunctive, prescriptive versus proscriptive).  
The focus on the formation of new social norms during times of uncertainty
has implications for interventions in social behavioral outcomes, 
research on cultural change, and the spread of cognitive associations. 
Moreover, \hyperlink{paper-3}{Article 3} focuses on how conflicting
behaviors become settled into cognitive associations, providing
a relevant framework for research on unrest and disasters.

@@

% social epidemiology
I also make important contributions to social epidemiology through this
research with the use of health case studies for questions of sociological interest
across my three articles. Because of this, I 
contribute to the social epidemiological conversation on vaccine hesitancy,
health communication, and diffusion of high-risk health behaviors. While
all these areas are important to social epidemiology, there hold
potential important health implications for these contributions on
outside of the academy, covered in the first paragraphs of this chapter.
